Measurement of total hadronic differential cross sections in the LArIAT experiment

### 2018-04-11 Chapter-1/1-Theory.tex
s

\begin{sidewaystable}
%\Rotatebox{90}{%
\begin{tabular}{| c | l | m{7cm} | m{8cm} |}
\hline
                    &              &                     &           \\
N $\pi$ in FS & Channel Name & Reaction &  Notes \\
                    &              &                     &           \\
\hline
                    &              &                     &           \\
 0  & \textbf{Pion Absorption}, $\sigma_{abs}$ &  \begin{tabular}{@{}l@{}} $\pi^- (np) \rightarrow nn$ (2-body abs) \\ $\pi^- (nnp) \rightarrow nnn$ (3-body abs)  \\ $\pi^- (npp) \rightarrow pnn$ (3-body abs)\\  $\pi^- (nnpp) \rightarrow pnn$ (Multi-body abs)  \end{tabular}  &Suppressed on single nucleon by energy conservation: the process occurs on at least two nucleons system.     \\
                    &              &                     &           \\
 \hline
                     &              &                     &           \\
 1 &  \textbf{Elastic Scattering}, $\sigma_{el}$ & $\pi^- + N \rightarrow \pi^- + N$   & Scattering on nucleon or nucleus, the target is left in ground state         \\
                    &              &                     &           \\
 1 &  \textbf{Charge Exchange}, $\sigma_{chex}$ & \begin{tabular}{@{}l@{}} $\pi^- + p \rightarrow \Delta^0 \rightarrow \pi^0+n$ \\ $\pi^- + N \rightarrow \pi^+ + $ nucleons  \end{tabular} &\begin{tabular}{@{}m{8cm}@{}}Single charge exchange: charged pion converts into neutral pion \\  \\Double charge exchange: charged pion converts into opposite charge pion  \end{tabular}   \\
                    &              &                     &           \\
  1 &  \textbf{Inelastic Scattering}, $\sigma_{inel}$ &    \begin{tabular}{@{}l@{}} $\pi^- + p \rightarrow \Delta^0 \rightarrow \pi^-+p$ (knock-out)  \\  $\pi^- + n \rightarrow \Delta^- \rightarrow \pi^-+n$ (knock-out)   \end{tabular}  &\begin{tabular}{@{}m{8cm}@{}} Other possible reactions: \\ Pure  Inelastic scattering:  population of low energy bound excited states \\ Nuclear break-up with nucleons or fragments knock-out    \end{tabular}   \\
                    &              &                     &           \\
  \hline
                      &              &                     &           \\
 2+ & \textbf{Pion Production}, $\sigma_{\pi prod}$    &           $\pi^- + N \rightarrow \ge2 \pi $ + nucleons            &  Possible if pion K.E $\ge$ 500 Mev/c       \\
                    &              &                     &           \\
\hline
\end{tabular}
%}
\caption{Summary of negative pion hadronic interactions as a function of the pion multiplicity in the final state in the energy range [100-1200] MeV. }
\label{tab:pionFinal}
\end{sidewaystable}

\subsubsection{Previous measurements: Lighter and Heavier Nuclei}
Many experiments with pion beams have studied the hadronic interaction of pions on light and heavy materials, such as  He, Li, C, Fe, Pb\textcolor{red}{\cite{s}}. However, data on argon are rare: the total differential hadronic cross section has never been measured before on argon. Simulation packages like Geant base their pion transportation for argon on data from lighter and heavier nuclei: the goal of LArIAT's dedicated measurement on argon is to bridge this gap in data, thus reducing the uncertainties related to pions interaction in argon in both neutrino event generators and in simulation packages of pion transportation.

The shape of the pion-nucleus interaction cross section in the energy range considered shows the distinct features that indicate the presence of a resonance. The mean free path of a pion of kinetic energy between 100 and 400 MeV is much shorter than the average distance between nucleons (which is of the order of 1 fm). Therefore, the pion interacts with surface nucleons. A $\Delta$ resonance is often produced in the interaction, which subsequently decays inside the nucleus.

Experimental results on several nuclei are reported in Figure as from \cite{PhysRevC.14.635}.

\begin{figure}[hbpt]
\centering
\includegraphics[]{Chapter-1/Images/PionsOnNuclei.png}
\caption{Pion-nucleus total cross sections:  $\sigma_{\pi^+}$ for positive pions (right) and $\sigma_{AV}$ (left) for the average between positive and negative pions $\sigma_{AV} = \frac{\sigma_{\pi^+} + \sigma_{\pi^-}}{2}$  in  the $\Delta$ resonance region.  The error bars include estimates of systematic uncertainties. The curves are the results of fits to the data assuming a Breit-Wigner shape. This summary plot  is reported in \cite{PhysRevC.14.635} and uses data from \cite{Clough1974, Wilkin:1973xd}.
}
\end{figure}

\subsubsection{Pion Interaction Cross Section for thin target in Geant4}

@@

\begin{figure}
\captionsetup{justification=raggedright}  
  \centering  
\includegraphics[width=3in]{Chapter-1/Images/CarbonG4.png}
\caption{total hadronic cross section for carbon implemented in Geant4  10.01.p3  with overlaid with the Bugg and Frideman data.}
\label{fig:TrueCarbon}
\end{figure}

>>>>

\begin{sidewaystable}[hp]
%\Rotatebox{90}{%
\begin{tabular}{| c | l | m{7cm} | m{8cm} |}
\hline
                    &              &                     &           \\
N $\pi$ in FS & Channel Name & Reaction &  Notes \\
                    &              &                     &           \\
\hline
                    &              &                     &           \\
 0  & \textbf{Pion Absorption}, $\sigma_{abs}$ &  \begin{tabular}{@{}l@{}} $\pi^- (np) \rightarrow nn$ (2-body abs) \\ $\pi^- (nnp) \rightarrow nnn$ (3-body abs)  \\ $\pi^- (npp) \rightarrow pnn$ (3-body abs)\\  $\pi^- (nnpp) \rightarrow pnn$ (Multi-body abs)  \end{tabular}  &Suppressed on single nucleon by energy conservation: the process occurs on at least two nucleons system.     \\
                    &              &                     &           \\
 \hline
                     &              &                     &           \\
 1 &  \textbf{Elastic Scattering}, $\sigma_{el}$ & $\pi^- + N \rightarrow \pi^- + N$   & Scattering on nucleon or nucleus, the target is left in ground state         \\
                    &              &                     &           \\
 1 &  \textbf{Charge Exchange}, $\sigma_{chex}$ & \begin{tabular}{@{}l@{}} $\pi^- + p \rightarrow \Delta^0 \rightarrow \pi^0+n$ \\ $\pi^- + N \rightarrow \pi^+ + $ nucleons  \end{tabular} &\begin{tabular}{@{}m{8cm}@{}}Single charge exchange: charged pion converts into neutral pion \\  \\Double charge exchange: charged pion converts into opposite charge pion  \end{tabular}   \\
                    &              &                     &           \\
  1 &  \textbf{Inelastic Scattering}, $\sigma_{inel}$ &    \begin{tabular}{@{}l@{}} $\pi^- + p \rightarrow \Delta^0 \rightarrow \pi^-+p$ (knock-out)  \\  $\pi^- + n \rightarrow \Delta^- \rightarrow \pi^-+n$ (knock-out)   \end{tabular}  &\begin{tabular}{@{}m{8cm}@{}} Other possible reactions: \\ Pure  Inelastic scattering:  population of low energy bound excited states \\ Nuclear break-up with nucleons or fragments knock-out    \end{tabular}   \\
                    &              &                     &           \\
  \hline
                      &              &                     &           \\
 2+ & \textbf{Pion Production}, $\sigma_{\pi prod}$    &           $\pi^- + N \rightarrow \ge2 \pi $ + nucleons            &  Possible if pion K.E $\ge$ 500 Mev/c       \\
                    &              &                     &           \\
\hline
\end{tabular}
%}
\caption{Summary of negative pion hadronic interactions as a function of the pion multiplicity in the final state in the energy range [100-1200] MeV. }
\label{tab:pionFinal}
\end{sidewaystable}

\subsubsection{Previous measurements: Lighter and Heavier Nuclei}
Many experiments with pion beams have studied the hadronic interaction of pions on light and heavy materials, such as  He, Li, C, Fe, Pb\textcolor{red}{\cite{s}}. However, data on argon are rare: the total differential hadronic cross section has never been measured before on argon. Simulation packages like Geant4 base their pion transportation for argon on data from lighter and heavier nuclei: the goal of LArIAT's dedicated measurement on argon is to bridge this gap in data, thus reducing the uncertainties related to pions interaction in argon in both neutrino event generators and in simulation packages of pion transportation.

The shape of the pion-nucleus interaction cross section in the energy range considered shows the distinct features that indicate the presence of a resonance. In fact, the mean free path of a pion of kinetic energy between 100 and 400 MeV is much shorter than the average distance between nucleons (which is of the order of 1 fm). Therefore, the pion interacts with surface nucleons. A $\Delta$ resonance is often produced in the interaction, which subsequently decays inside the nucleus.
Experimental results on several nuclei as reported in \cite{PhysRevC.14.635} are shown in Figure \ref{fig:pi1}; it is interesting to notice here how the shape fo the $\Delta$ resonance become less pronounced as a function of the mass number of the target nucleus. Pion interactions with heavier nuclei also shift the peak of the resonance at lower energy; this effect is due to kinematic considerations and to the difference in propagation of the $\Delta$ inside the nucleus. Multiple scattering effect modify the resonance width, which is larger than the natural-decay width.
As an example of a fairly well studied target, Figure \ref{fig:pi2} reports the negative pion cross section on Carbon for the elastic and reaction\footnote{What we refer as to ``reaction channel" is called ``inelastic" in the paper.} channels, and their sum \cite{BINON1970168}.

\begin{figure}[hbpt]
\centering
\includegraphics[scale=0.7]{Chapter-1/Images/PionsOnNuclei.png}
\caption{Pion-nucleus total cross sections:  $\sigma_{\pi^+}$ for positive pions (right) and $\sigma_{AV}$ (left) for the average between positive and negative pions $\sigma_{AV} = \frac{\sigma_{\pi^+} + \sigma_{\pi^-}}{2}$  in  the $\Delta$ resonance region.  The error bars include estimates of systematic uncertainties. The curves are the results of fits to the data assuming a Breit-Wigner shape. This summary plot  is reported in \cite{PhysRevC.14.635} and uses data from \cite{Clough1974, Wilkin:1973xd}.
}
\label{fig:pi1}
\end{figure}

\begin{figure}[hbpt]
\centering
\includegraphics[scale=0.7]{Chapter-1/Images/PionCarbon.png}
\caption{Negative pion nucleus total, elastic and reaction cross sections on $^{12}$C as from \cite{BINON1970168}.  }
\label{fig:pi2}
\end{figure}

\subsubsection{Pion Interaction Cross Section in generators}
A broad review of the current status of negative and positive pion simulation in Geant4 and GENIE for $^{12}$C, $^{56}$Fe, and $^{40}$Ca can be found in \cite{Irene} \textcolor{red}{Irene}. From that work, we report the results for $^{12}$C (left) in Figure \ref{fig:pionC} and  $^{40}$Ca (right) in Figure \ref{fig:pionCa}, since the argon cross section should lie in between.

\begin{figure}
\captionsetup{justification=raggedright}  
	\begin{minipage}[t]{.53\textwidth}  
	  \centering  
	   \includegraphics[width=3in]{Chapter-1/Images/IreneCarbon.png}
	   	        \caption{ Total, elastic and reaction cross section for $\pi^-$ on $^{12}$C. Comparison between results from Geant4 simulation (Bertini cascade model) and experimental data \cite{Clough1974, PhysRevC.23.2173, ALLARDYCE19731,osti_4488432}.}
        \label{fig:pionC}
	\end{minipage}%  
	\begin{minipage}[t]{0.53\textwidth}  
	  \centering  
	\includegraphics[width=3in]{Chapter-1/Images/IreneCalcium.png}
        \caption{Total, elastic and reaction cross section for $\pi^-$ on $^{40}$Ca. Comparison between results from Geant4 simulation (Bertini cascade model),  Genie simulation (hA model), and experimental data \cite{ALLARDYCE19731}.}
        \label{fig:pionCa}
	\end{minipage}
	\par
\end{figure}

@@

\begin{figure}[b!]
\captionsetup{justification=raggedright}  
  \centering  
\includegraphics[width=3in]{Chapter-1/Images/CarbonG4.png}
\caption{total hadronic cross section for carbon implemented in Geant4  10.01.p3  with overlaid with the Bugg and Frideman data.}
\label{fig:TrueCarbon}
\end{figure}
